---
title: "Equipment Energy Composition"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
---

100% stacked bar chart of BL equipment energy composition, for the main text: each
segment is one equipment type's share of total energy use, computed as the ratio of
averages - mean `annual_electricity_<type>` across labs, divided by the sum of all
types' means (which matches `annual_electricity_total`'s mean exactly - the 11 types
are exhaustive, no residual "other" category needed). Segments are ordered 
largest-to-smallest by the "overall" bar, that order is kept fixed across the 
by-faculty bars too, so the same equipment type lines up in the same position.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.show = "show")
```

```{r config-and-libraries, include=FALSE}
library(here)
library(dplyr)

# Load config and functions
source(here("config.R"))
source(here("3_Analysis", "functions", "stacked_bar_energy_share_func.r"))

# Output directory
hist_dir <- file.path(OUTPUT, "6_Histograms")
dir.create(hist_dir, showWarnings = FALSE, recursive = TRUE)
```

```{r load-data}
df <- read.csv(file.path(CLEAN_DATA, "final_dataset.csv"))
```

```{r restrict-sample}
# Same "BL, sample with matched EL data" restriction as the electricity histograms
df_sample <- df %>%
  group_by(labgroupid) %>%
  filter(n() == 2) %>%
  ungroup()

df_bl <- df_sample %>% filter(survey == "BL") %>%
  mutate(MNF = ifelse(faculty == "Faculty of Science (MNF)", 1, 0))
```

```{r energy-cols}
# Column -> display label, for all 11 equipment types that sum to
# annual_electricity_total (matches the labelling used in
# 3_8_hist_equipment.rmd's equipment counts)
energy_cols <- c(
  annual_electricity_it        = "IT Equipment",
  annual_electricity_fc        = "Fume Cupboards",
  annual_electricity_freezer   = "Freezers",
  annual_electricity_ult       = "ULT Freezers",
  annual_electricity_fridge    = "Fridges",
  annual_electricity_incubator = "CO2 Incubators",
  annual_electricity_microbio  = "MSCs",
  annual_electricity_glassware = "Drying Cabinets",
  annual_electricity_bath      = "Water Baths",
  annual_electricity_cryostat  = "Cryostats",
  annual_electricity_heater    = "Block Heaters"
)
```

```{r equipment-share-bar}
make_energy_share_bar(
  df           = df_bl,
  energy_cols  = energy_cols,
  bar_var      = "MNF",
  bar_labels   = list("0" = "Medicine or Joint", "1" = "Science"),
  output_file  = "equipment_energy_share.pdf",
  font_family  = "serif",  # matches the paper's Times-style LaTeX typesetting
  output_dir   = hist_dir
)
```
