---
title: "Electricity Consumption Histograms (Sensitivity-Proof, Truncated)"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: true
---

This is a sensitivity-proof variant of `3_1_hist_electricity.rmd`, for any audience
where the plain histogram there is too granular: `annual_electricity_total` is
strongly right-skewed with a handful of labs far out in the tail (e.g. single
labs alone in a 10,000 kWh-wide bin above 100,000 kWh). Two fixes,
applied together via `make_truncated_histogram()`:

1. **Open-ended top bin.** Everything at or above `THRESHOLD` (50 MWh) is
   collapsed into a single bar, labelled `>=50`, showing only the
   aggregate share of labs up there - never a finer breakdown of who they are.
2. **Minimum bin-width rule.** Below `THRESHOLD`, the binwidth is chosen so
   every bin holds at least `MIN_BIN_COUNT` labs, checked directly against the
   data for every panel below. The overall and BL-vs-EL panels have enough
   labs to support a finer `OVERALL_BINWIDTH`; the treatment/faculty splits
   divide the sample further and need the wider `GROUPED_BINWIDTH` to still
   satisfy the rule in their smallest subgroup. The function warns if a chosen
   binwidth doesn't hold for a given panel.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.show = "show")
```

```{r config-and-libraries, include=FALSE}
library(here)
library(dplyr)

# Load config and functions
source(here("config.R"))
source(here("3_Analysis", "functions", "histogram_truncated_func.r"))

# Output directory
hist_dir <- file.path(OUTPUT, "6_Histograms")
dir.create(hist_dir, showWarnings = FALSE, recursive = TRUE)
```

```{r load-data}
df <- read.csv(file.path(CLEAN_DATA, "final_dataset.csv"))
```

```{r create-vars}
# Create MNF variable
df <- df %>% mutate(MNF = ifelse(faculty == "Faculty of Science (MNF)", 1, 0))

# Work in MWh rather than kWh - more readable bin labels once the axis is a
# small number of discrete categories rather than a continuous kWh scale
df <- df %>% mutate(annual_electricity_mwh = annual_electricity_total / 1000)
```

```{r separate-bl-el}
# Restrict df to only those labgroupids with BL and EL data
df_sample <- df %>%
  group_by(labgroupid) %>%
  filter(n() == 2) %>%
  ungroup()

# Separate BL and EL
df_bl_all <- df %>% filter(survey == "BL")
df_bl_sample <- df_sample %>% filter(survey == "BL")
df_el <- df %>% filter(survey == "EL")
```

```{r disclosure-params}
# Truncation threshold and minimum bin-width rule, applied consistently across
# every panel below. Two binwidths, both the finest that still satisfies
# "every bin >= MIN_BIN_COUNT labs" for their respective panels (checked
# against the data, not assumed):
#   - OVERALL_BINWIDTH: for the overall and BL-vs-EL panels, which have
#     enough labs to support finer resolution than the grouped panels.
#   - GROUPED_BINWIDTH: for the treatment/faculty splits, which divide the
#     sample further - their smallest subgroup (EL, non-MNF, n=32) needs the
#     wider bin to still hold >= MIN_BIN_COUNT labs per bin.
THRESHOLD        <- 50    # MWh
OVERALL_BINWIDTH <- 12.5  # MWh
GROUPED_BINWIDTH <- 25    # MWh
MIN_BIN_COUNT    <- 5
LABEL_SUFFIX     <- " MWh"   # unit folded into each bin label, e.g. "0-12.5 MWh"
FONT_FAMILY      <- "serif"  # matches the paper's Times-style LaTeX typesetting
```

# Overall Distribution

```{r overall-bl}
# BL overall - all groups
make_truncated_histogram(
  df            = df_bl_all,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = OVERALL_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  output_file   = "hist_electricity_bl_overall_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```

```{r overall-bl-sample}
# BL overall - sample with BL and EL data
make_truncated_histogram(
  df            = df_bl_sample,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = OVERALL_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  output_file   = "hist_electricity_bl_overall_sample_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```

```{r overall-el}
# EL overall
make_truncated_histogram(
  df            = df_el,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = OVERALL_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  output_file   = "hist_electricity_el_overall_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```


# By Treatment Status

```{r treatment-bl}
# BL by treatment - all groups
make_truncated_histogram(
  df            = df_bl_all,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = GROUPED_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  group_var     = "treated",
  group_labels  = list("0" = "Control", "1" = "Treatment"),
  output_file   = "hist_electricity_bl_by_treatment_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```

```{r treatment-bl-sample}
# BL by treatment - sample with BL and EL data
make_truncated_histogram(
  df            = df_bl_sample,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = GROUPED_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  group_var     = "treated",
  group_labels  = list("0" = "Control", "1" = "Treatment"),
  output_file   = "hist_electricity_bl_by_treatment_sample_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```

```{r treatment-el}
# EL by treatment
make_truncated_histogram(
  df            = df_el,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = GROUPED_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  group_var     = "treated",
  group_labels  = list("0" = "Control", "1" = "Treatment"),
  output_file   = "hist_electricity_el_by_treatment_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```

# By Faculty

```{r faculty-bl}
# BL by faculty - all groups
make_truncated_histogram(
  df            = df_bl_all,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = GROUPED_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  group_var     = "MNF",
  group_labels  = list("0" = "Medicine or Joint", "1" = "Science"),
  output_file   = "hist_electricity_bl_by_faculty_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```

```{r faculty-bl-sample}
# BL by faculty - sample with BL and EL data
make_truncated_histogram(
  df            = df_bl_sample,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = GROUPED_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  group_var     = "MNF",
  group_labels  = list("0" = "Medicine or Joint", "1" = "Science"),
  output_file   = "hist_electricity_bl_by_faculty_sample_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```

```{r faculty-el}
# EL by faculty
make_truncated_histogram(
  df            = df_el,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = GROUPED_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  group_var     = "MNF",
  group_labels  = list("0" = "Medicine or Joint", "1" = "Science"),
  output_file   = "hist_electricity_el_by_faculty_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```


# BL vs EL Comparison

```{r bl-vs-el}
# BL vs EL comparison - sample with BL and EL data
make_truncated_histogram(
  df            = df_sample,
  variable      = "annual_electricity_mwh",
  threshold     = THRESHOLD,
  binwidth      = OVERALL_BINWIDTH,
  min_bin_count = MIN_BIN_COUNT,
  group_var     = "survey",
  output_file   = "hist_electricity_bl_vs_el_truncated.pdf",
  show_median   = TRUE,
  median_digits = 1,
  label_suffix  = LABEL_SUFFIX,
  font_family   = FONT_FAMILY,
  output_dir    = hist_dir
)
```
